---
title: "HCUP NY 2017-2018 EDA"
author: "Ran Li"
date: "12/08/2022"
format: 
  html:
    toc: true
    self-contained: true
editor: visual
editor_options: 
  chunk_output_type: console
execute: 
  warning: false
  error: false  
  echo: false
---

## Overview

```{r}
#| echo: false

pivot_vars = function(df){
  df %>% 
    add_count(var) %>% 
    arrange(desc(n), var) %>% 
    select(var,  dataset_id) %>% 
    mutate(value = T) %>% 
    pivot_wider(names_from = var, values_from = value)
}

```

Some exploratory data analysis just on the HCUP metadata. Below

```{r}
library(tidyverse)
library(reactable)
df_codebooks = read_csv("../../clean/df_codebooks.csv") %>% 
  rowwise() %>% 
  mutate(split = str_split(dataset_id,pattern = "_"),
         state = split[[1]],
         database = split[[2]],
         year = split[[3]],
         dataset = split[[4]]
         )  %>% 
  select(dataset_id, var, var_label, value_type, everything(), - split)

## Summarize number of variables
df_codebooks %>% 
  count(database,dataset, year, name = 'n_vars') %>% 
  mutate(year  = paste0("n_var_",year) ) %>% 
  # mutate(value = T) %>% 
  pivot_wider(names_from = year, values_from = n_vars) %>% 
  reactable()

```


- **Number of variables in CORE datasets change by year. Increasing a lot particularly for SEDD from 2017 to 2018.**
- CHGS datasets don't change much.


Lets look into these more closely below


## CHGS


```{r}
df_codebooks %>% 
  filter(dataset == "CHGS") %>% 
  pivot_vars()%>% 
  reactable()
```


- 4 CHGS fields for both SEDD and SID:
- 4 CHGS fields unique to CHGS
- These remain constant between 2017 and 2018

## SEDD CORE

```{r}
## Table of vars by year 
df_sedd_core = df_codebooks %>% 
  filter( database=="SEDD", dataset == "CORE") %>% 
  arrange(var) %>% 
  select(var, year) %>% 
  mutate(value = T) %>% 
  pivot_wider(names_from = year, values_from = value)


## Summary of vars in years
df_sedd_core %>% 
  mutate(var_short = str_remove(var, "[0-9]*$"),
         var_appendix = str_extract(var,"[0-9]*$")) %>%
  add_count(var_short, name = 'n_var') %>%   
  group_by(var_short,n_var, `2017`, `2018`) %>% 
  summarize(var_appendix = paste(var_appendix, collapse = ", ") %>% str_sub(1,15)) %>% 
  ungroup() %>% 
  arrange(desc(n_var)) %>% 
  reactable()

```

- The SEDD Core table is huge with 603 columns in 2017 and 1106 colums in 2018.
- The majority of these columns are some type of diagnosis codes including:
   - **CPT**###: CPT/HCPCS procedure code
   - **COTCSS**###: The CCS-Services and Procedures provides a method for classifying Healthcare Common Procedure Coding System (HCPCS) codes into clinically meaningful categories.
   - **CPTDAY**###: Number of days from admission to CPT ###
   - **CPTM1_**###: First CPT-4/HCPCS modifier ###
   - **CPTM1_**###: Second CPT-4/HCPCS modifier ###
   - **I10_DX**##: 	ICD-10-CM Diagnosis ##
- In 2018 there are a lot of new codes for the above column types; these account for most of the changes in variables between years
- There are also small changes in other variables such as `MARITALSTATUS_X` or `PointOfOriginUB`
   

#### SID CORE

```{r}
## Table of vars by year 
df_sid_core = df_codebooks %>% 
  filter( database=="SID", dataset == "CORE") %>% 
  arrange(var) %>% 
  select(var, year) %>% 
  mutate(value = T) %>% 
  pivot_wider(names_from = year, values_from = value)


## Summary of vars in years
df_sid_core %>% 
  mutate(var_short = str_remove(var, "[0-9]*$"),
         var_appendix = str_extract(var,"[0-9]*$")) %>%
  add_count(var_short, name = 'n_var') %>%   
  group_by(var_short,n_var, `2017`, `2018`) %>% 
  summarize(var_appendix = paste(var_appendix, collapse = ", ") %>% str_sub(1,15)) %>% 
  ungroup() %>% 
  arrange(desc(n_var)) %>% 
  reactable()

```

- The SID Core table is large with 251 columns in 2017 and 255 columns in 2018.
- The majority of these columns are some type of diagnosis codes including:
   - **DXPOA**###: Diagnosis ##, present on admission indicator
   - **I10_DX**##: ICD-10-CM Diagnosis ##
   - **I10_PR**###: ICD-10-PCS Procedure  ##
   - **PRDAY**###: Number of days from admission to I10_PR##
   - **PRMONTH**###: Month of procedure ##
   - **PRYEAR**###: Year of procedure ##
- In 2018 there only four news codes for the above column types; these account for most of the changes in variables between years
